library(tidyverse)
library(readr)
library(leaflet)

Suicide_Mortality_by_State <- read_csv("Suicide Mortality by State.csv")

region <- read_csv("region.csv")
```

```{r}
region_suicide <- region %>% 
  rename(STATE= "State Code") %>% 
  full_join(Suicide_Mortality_by_State, by="STATE") %>% 
  filter(YEAR!= "2005")
```

```{r}
not_2005 <- Suicide_Mortality_by_State %>% 
  filter(YEAR!= "2005")

ggplot(data=not_2005, mapping=aes(group= YEAR, x=YEAR, y=RATE))+
  geom_boxplot()+
  labs(x= "Year", y= "Rate", title= "Suicide Deaths per 100,000 People")+
  scale_x_continuous( breaks=seq(2014,2017,1))

ggplot(data=region_suicide, mapping=aes(x=YEAR, y=RATE, group= YEAR))+
  geom_boxplot()+
  scale_x_continuous( breaks=seq(2014,2017,1))+
  facet_wrap(~Region)
  ```
  
  
 ```{r}
 #wrangling income
income<-kaggle_income %>% 
  select(Median,State_Name) %>% 
  filter(!State_Name == "Puerto Rico", !State_Name == "District of Columbia") %>% 
  group_by(State_Name) %>%  
  summarize(median_income = median(Median))

View(income)
 #wrangling location
location <-statelatlong_1_ %>% 
  filter(!City == "Puerto Rico", !City == "District of Columbia") 

View(location)

#combining location and income
income_location <- income %>% 
  full_join(location, by = c("State_Name" = "City" )) 

View(income_location)
```
