# FIRST PAGE

---
title: "Suicide Rates in the United States (2014-2017)"
author: "Naomi Giancola, Christina Sherpa, Tessa Finkelstein"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: kable
---

<!-- 
Comments in HTML are like this! 
-->

# Suicide Deaths per 100,000 People

Write an intro `index.html` page in blog-post style that summarizes your work. It should have

1. One visualization that best summarizes your work.
1. No more than 200 words of text

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(leaflet)

MH_12 <- read_csv("MH_12.csv")

Suicide_Mortality_by_State <- read_csv("Suicide Mortality by State.csv")

region <- read_csv("region.csv")
```

```{r, echo=FALSE}
not_2005 <- Suicide_Mortality_by_State %>% 
  filter(YEAR!= "2005")

region_suicide <- region %>% 
  rename(STATE= "State Code") %>% 
  full_join(Suicide_Mortality_by_State, by="STATE") %>% 
  filter(YEAR!= "2005")

ggplot(data=not_2005, mapping=aes(group= YEAR, x=YEAR, y=RATE))+
  geom_boxplot()+
  labs(x= "Year", y= "Rate", title= "Suicide Deaths per 100,000 People in the United States (by Year)")+
  scale_x_continuous( breaks=seq(2014,2017,1))+
  theme_dark()
  ```
  
# SECOND PAGE

---
title: "More information"
author: "Naomi Giancola, Christina Sherpa, Tessa Finkelstein"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: kable
---

<!--
Comments in HTML are like this! 
-->

# Suicide by Region in the United States

Write a second `more_info.html` page going more in-depth for people who want more details. In should have

1. Between 2-3 more visualizations.
1. No more than 500 words of text

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(leaflet)
MH_12 <- read_csv("MH_12.csv")

Suicide_Mortality_by_State <- read_csv("Suicide Mortality by State.csv")

region <- read_csv("region.csv")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
not_2005 <- Suicide_Mortality_by_State %>% 
  filter(YEAR!= "2005")

region_suicide <- region %>% 
  rename(STATE= "State Code") %>% 
  full_join(Suicide_Mortality_by_State, by="STATE") %>% 
  filter(YEAR!= "2005")
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(data=region_suicide, mapping=aes(x=YEAR, y=RATE, group= YEAR))+
  geom_boxplot()+
  scale_x_continuous( breaks=seq(2014,2017,1))+
  facet_wrap(~Region)+
  labs(x= "Year", y= "Rate", title= "Suicide Deaths per 100,000 People (by Region)")+
  theme_dark()

ggplot(data = region_suicide, mapping = aes(x = RATE, y = DEATHS)) +
  geom_point(aes(color = Region)) +
  facet_wrap(~YEAR) +
  labs(x= "Rate", y= "Deaths", title= "Suicide Attempt Rates vs. Actual Deaths (by Region)") +
  scale_color_manual(values = c("#f1eef6", "#bdc9e1", "#74a9cf", "#0570b0")) +
  theme_dark()

ggplot(data=region_suicide, mapping=aes(x=YEAR, y=RATE, fill=Division))+
  geom_col(position = "dodge")+
  labs(x= "Year", y= "Rate", title= "Suicide Deaths per 100,000 People (by Division)")+
  scale_fill_manual(values = c("#fff7fb", "#ece7f2", "#d0d1e6", "#a6bddb", "#74a9cf", "#3690c0", "#0570b0", "#045a8d", "#023858"))+
  theme_dark()
```
